R Notebook

### 2022-02-19 docs/3. SNA - Assignments/sna_Week3/Week3_Notebook.Rmd
11.??.22
I took a break

# Network Info
  Network Objects: EDGE LIST
                  STATNET: smugEL_sn
                  IGRAPH: smugEL_ig

```{r EDGELIST: Network Objects: statnet & igraph, include=FALSE}

@@

@@

```{r igraph: Network Description, echo=TRUE}

      # class of an object
    class(smugEL_ig )

@@

## igraph: Exploring Network

```{r}
library("igraph")

```
IGRAPH 574c3d4 UN-- 274 672 -- 
+ attr: name (v/c), 
        dnkName (v/c), 
        dnkIng (e/c),
      | ingType (e/c)
+ edges from 574c3d4 (vertex names):

#   Week 2: Basic Network Structure: Smuggler's Cove Drink Index
We can break large social networks down into their constituent parts. These constituent parts are referred to as “motifs”. The most basic motif consists of two nodes and is called a dyad. Edges in a network signify the presence or absence of dyadic relations. It follows that a dyad in an undirected network can have two unique configurations: connected or disconnected; and three unique configurations in a directed network (mutual, assymetric, and null)

@@

```{r iGraph:Network: Path: Shortest, echo=TRUE}

@@

```{r iGraph: Specific Nodes: Path: Shortest, echo=TRUE}

@@

```{r iGraph: Specific Nodes: Path: Length, echo=TRUE}

@@

@@

```{r iGraph: Dyad Census, echo=TRUE}

graph.density(smugEL_ig)
    ## Density captures, at the macro-level, the proportion of dyads that are present over the possible total number of dyads in the network

@@

@@

```{r iGraph: Triad Census, echo=TRUE}
#Classify all Triad in the network
     igraph::triad_census(smugEL_ig)

@@

```{r iGraph: weighted/ network transitivity, echo=TRUE}

@@

## Local Transivity or Clustering
`Local transitivity` (local clustering coefficient), is a technical description of the density of an ego network.
```{r iGraph: Transitivity: Network}
#     get global clustering cofficient
  transitivity(smugEL_ig, type="global")

@@

```{r iGraph:Local transitivity: Node List & Ego Network, echo=TRUE}

@@

```{r igraph: Component Structure, echo=TRUE}
                ## Component Structure:

@@

```{r igraph: network density, echo=TRUE}

@@

```{r igraph: Vertex Degree, echo=TRUE}
# igraph: Calculate average network degree
 # provides node name along with node degree

@@

```{r Statnet: , echo=TRUE}
      # Basic R: get summary statistics for node attributes
 summary(network_nodes)
      # provides range, mean, median and standard deviation of a vector.

@@

#Visualizations

## igraph Visualizations
igraph provides three different ways for visualization.
```{r igraph: , echo=TRUE}

plot.igraph(smugEL_ig)

 # uses regular R graphics and can be used with any R device.

#  (Tk is quite resource hungry, so don't try this for very large graphs.)

 tkplot(smugEL_ig)

@@

Distill is a publication format for scientific
and technical writing, native to the web.
>
Learn more about using Distill for R Markdown at
<https://rstudio.github.io/distill>. ##

>>>>

# Network Info
  Network Objects: EDGE LIST
                  STATNET: smugEL_sn
                  IGRAPH: smugEL_ig
                          ingTypeEL_Dig

```{r EDGELIST: UNDIRECTED Network Objects: statnet & igraph, include=FALSE}

@@

```{r EDGELIST: IGRAPH: DIRECTED Network: ingType, include=FALSE}

  nodes_ingType <- smugCovIdx 
    nodes_ingType<- 
      select(nodes_ingType,ingType)
view(nodes_ingType)
        ### View(nodes_ingType)

######
      ######**igraph** 
######
        #igraph object: ingType
d_network_ig <- 
  graph.edgelist(
    as.matrix(
      smugCovIdx[2:3]), 
    directed=T)

vertex.attributes(d_network_ig)
                ## create new edge attribute
  E(d_network_ig)$dnkIng<-smugCovIdx$dnkIng
  E(d_network_ig)$dnkName<-smugCovIdx$dnkName

  #create vertex attributes in igraph
 V(d_network_ig)[match(V(d_network_ig)$ingType,nodes_ingType$ingType)]$ingType<- nodes_ingType$ingType

   View(d_network_ig)
   head(d_network_ig)
  summary(d_network_ig)
  summary(edge_attr(d_network_ig))
  summary(vertex.attributes(d_network_ig))

   ingTypeEL_Dig <- d_network_ig
```

@@

```{r STATNET: functions that calculate network-level descriptives}
library(statnet)
library(sna)

    #summary of network object, also provides an edge list
summary(smugEL_sn)$no

    #number of nodes in the network
network.size(smugEL_sn)

    #number dyads that exist in network (n*(n-1))
network.dyadcount( smugEL_sn)

    #number of edges present in the network
network.edgecount( smugEL_sn)

    #network density (edge count/dyad count)
gden(smugEL_sn)
  gden(smugEL_sn,
       diag = T,
       mode = "graph")

    ##**diag=* 
    ##boolean indicating whether or not the diagonal should be treated as valid data. 
        ##*Set this true if and only if the data can contain loops. 
        ##*diag is FALSE by default.
    ##**mode="digraph"*
    ##* string indicating the type of graph being evaluated. 
    ##* "digraph" indicates that edges should be interpreted as directed; 
    ##* "graph" indicates that edges are undirected. mode is set to "digraph" by default.

  # grecip calculates the dyadic reciprocity 
grecip(smugEL_sn) 
      #proportion of symmetric dyads
  grecip(smugEL_sn, measure = "dyadic") 
      # returns the ratio of mutuals to non-null dyads.
  grecip(smugEL_sn, measure = "dyadic.nonnull") 
      # the proportion of edges which are reciprocated
  grecip(smugEL_sn, measure = "edgewise") 
      # returns the log of the ratio of the edgewise reciprocity to the density
  grecip(smugEL_sn, measure = "edgewise.lrr") 
      # The correlation measure may be interpreted as the net tendency for edges of similar relative value (with respect to the mean edge value) to occur within the same dyads. 
  grecip(smugEL_sn, measure = "correlation")

      # k-path or k-cycle  compute census statistics (respectively) on one or more input graphs
      ## In addition to aggregate counts of paths or cycles, results may be disaggregated by vertex and co-membership information may be computed.

Usage

  #**kpath* Count paths of length <=3
kpath.census(smugEL_sn, maxlen=3,
             tabulate.by.vertex=FALSE) 
View(kpath.census(smugEL_sn, tabulate.by.vertex=T))

#**kcycle*Count cycles of length <=3
kcycle.census(smugEL_sn, maxlen=3, tabulate.by.vertex=FALSE) 
  View(kcycle.census(smugEL_sn, maxlen=3,tabulate.by.vertex=T))

# counts of cliques by size
clique.census(smugEL_sn, tabulate.by.vertex=F, enumerate=F)
view(clique.census(smugEL_sn, tabulate.by.vertex=TRUE, enumerate=TRUE))

gtrans(smugEL_sn)
proportion of transitive triads

symmetrize( smugEL_sn, rule = "weak") 
symmetrize so i<->j iff i->j OR i<-j

symmetrize( smugEL_sn, rule = "strong") 
symmetrize so i<->j iff i->j OR i<-j

dyad.census(smugEL_sn) 
MAN dyad census

triad.census(smugEL_sn) 
standard directed triad census

```{r igraph: Network Description, echo=TRUE}
library(igraph)
      # class of an object
    class(smugEL_ig)

@@

```{r IGRAPH: DIRECTED Network: ingType, echo=TRUE}

      # class of an object
    class(ingTypeEL_Dig)

      # Count Vertices
    vcount(ingTypeEL_Dig)

      # Count Edges
    ecount(ingTypeEL_Dig)

## Network Mode
    # Bipartitie / Unimodal
    is_bipartite(ingTypeEL_Dig)

## Edges    
    # Un/directed 
    is_directed(ingTypeEL_Dig)

    # weighted or unweighted?
    is_weighted(ingTypeEL_Dig)

    # edge weights
    E(ingTypeEL_Dig)$weight

    # Generate a list the names of vertex attributes
  V(ingTypeEL_Dig)$name

    # access edge attribute  names
  igraph::vertex_attr_names(ingTypeEL_Dig)

    # Generate a list the names of edge attributes  
  igraph::edge_attr_names(ingTypeEL_Dig)

## Network Attributes
 summary(ingTypeEL_Dig)
          # Vertices =  
          # Directed = 
          # hyper = 
          # loops = 
          # multiple = 
          # bipartite = 
          # total edges = 
                # missing edges =
                # non-missing edges = 
          # Vertex attribute names:
          # Edge Attribute Names:

  # class of an object
[1] "igraph"

  # Count Vertices
[1] 201

  # Count Edges
[1] 672

      ## Network Mode
  # Bipartitie / Unimodal
     is_bipartite
[1] FALSE

    ## Edges    
   # Un/directed 
     is_directed
[1] TRUE

  # weighted or unweighted?
     is_weighted
[1] FALSE

   # edge weights
NULL

#   Week 2: Basic Network Structure: Smuggler's Cove Drink Index
)

@@

### Igraph: Undirected
```{r iGraph: Undirected: Path: Shortest, echo=TRUE}

@@

```{r iGraph: Undirected: Specific Nodes: Path: Shortest, echo=TRUE}

@@

```{r iGraph: Undirected: Specific Nodes: Path: Length, echo=TRUE}

@@

```
### Igraph: Directed
```{r iGraph: Directed: Path: Shortest, echo=TRUE}

#     find average shortest path for network
    average.path.length(ingTypeEL_Dig,directed=T)

```{r iGraph: Directed:Specific Nodes: Path: Shortest, echo=TRUE}

#     isolating the distances between specific node
    all_shortest_paths(ingTypeEL_Dig,
                       "Strozzi","Tornabuoni", 
                       weights=NA)$res
## Note: manually tell igraph to ignore edge weights.
```

```{r iGraph: Directed: Specific Nodes: Path: Length, echo=TRUE}

#     Calculate distances between two nodes
    distances(ingTypeEL_Dig,"Bischeri","Ridolfi")

#     Calculate distance between two nodes using unweighted edges
    distances(ingTypeEL_Dig,"Bischeri", "Castellani",weights=NA)

@@

```
plot(sna::dyad.census(smugEL_sn))
```{r iGraph: Undirected: GRAPH: Dyad Census, echo=TRUE}
library(igraph)
graph.density(smugEL_ig)
    ## Density captures, at the macro-level, the proportion of dyads that are present over the possible total number of dyads in the network

@@

```
plot(sna::triad.census(smugEL_sn, mode="graph"))
plot(sum(sna::triad.census(smugEL_sn, mode="graph")))

```{r iGraph: Undirected: Triad Census: Calculating a triad census, echo=TRUE}
#Classify all Triad in the network
   igraph::triad_census(smugEL_ig)

@@

 plot(igraph::triad_census(smugEL_ig))

```{r iGraph: Undirected:TRIDADS: Banning triads}

@@

```{r iGraph: Undirected: weighted/ network transitivity, echo=TRUE}

@@

## Local Transivity or Clustering
`Local transitivity` (local clustering coefficient), is a technical description of the density of an ego network.
```{r iGraph: Undirected: Transitivity: Network}
#     get global clustering cofficient
  transitivity(smugEL_ig, type="global")

@@

```{r iGraph: Undirected: Local transitivity: Node List & Ego Network, echo=TRUE}

@@

```{r iGraph: Undirected: Component Structure, echo=TRUE}
                ## Component Structure:

@@

```{r iGraph: Undirected: network density, echo=TRUE}

@@

```{r iGraph: Undirected: Vertex Degree, echo=TRUE}
# igraph: Calculate average network degree
 # provides node name along with node degree

@@

```{r Statnet: , echo=TRUE}
      # Basic R: get summary statistics for node attributes
 summary(nodes_ingTypes)
      # provides range, mean, median and standard deviation of a vector.

@@

# Centrilization

```{r iGraph: Undirected: DEGREE CENTRALITY: dnkIng}
library(igraph)
library(tidyverse)

# graph the SMUG network

smug_attributes <- smugCovIdx
smug_edj <- smugCovIdx[2:3]
View(smug_edj)
#V1: dnkIng
#V2: ingType

smugNet <- 
  graph.edgelist(as.matrix(smug_edj), directed = T)

V(smugNet)$dnkIng <- 
  smug_attributes$dnkIng[match(V(smugNet)$name, smug_attributes$ingType)]

V(smugNet)$dnkName <- 
  smug_attributes$dnkName[
    match(V(
      smugNet)$name,
      smug_attributes$dnkIng)
    ]
summary(smugNet)

###**PLOT** in .R Doc

```{r iGraph: Undirected: DEGREE CENTRALITY: dnkIng}
library(igraph)
library(tidyverse)

# graph the SMUG network

smug_attributes <- smugCovIdx
smug_edj <- smugCovIdx[2:3]
View(smug_edj)
#V1: dnkIng
#V2: ingType

smugNet <- 
  graph.edgelist(as.matrix(smug_edj), directed = T)

V(smugNet)$ingType <- 
  smug_attributes$ingType[match(V(smugNet)$name, smug_attributes$ingType)]

V(smugNet)$dnkName <-
  smug_attributes$dnkName[match(V(smugNet)$name, smug_attributes$dnkName)]

summary(smugNet)

###**PLOT** in .R Doc

```{r iGraph: Undirected: DEGREE CENTRALITY: ingType, echo=TRUE}
# graph the SMUG network

smug_attributes <- smugCovIdx
smug_edj <- smugCovIdx[2:3]
View(smug_edj)
#V1: dnkIng
#V2: ingType

smugNet <- 
  graph.edgelist(as.matrix(smug_edj), directed = T)

V(smugNet)$dnkName <-
  smug_attributes$dnkName[match(V(smugNet)$name, smug_attributes$dnkIng)]

summary(smugNet)
View(smugNet)

###**PLOT** in .R Doc

@@

Distill is a publication format for scientific
and technical writing, native to the web.

Learn more about using Distill for R Markdown at
<https://rstudio.github.io/distill. ##

# iGraph Directed Network Object

## Degree Centrality
We can break large social networks down into their constituent parts. These constituent parts are referred to as “motifs”. The most basic motif consists of two nodes and is called a dyad. Edges in a network signify the presence or absence of dyadic relations. It follows that a dyad in an undirected network can have two unique configurations: connected or disconnected; and three unique configurations in a directed network (mutual, assymetric, and null

```{r iGraph: Directed: GRAPH: Dyad Census, echo=TRUE}

graph.density(ingTypeEL_Dig)
    ## Density captures, at the macro-level, the proportion of dyads that are present over the possible total number of dyads in the network

ingTypeEL_DigNod <- vcount(ingTypeEL_Dig)

ingTypeEL_DigDen <- graph.density(ingTypeEL_Dig)

smugEL_DigRandoG <- 
  erdos.renyi.game(
    n = smugEL_igNod, 
    p.or.m = smugEL_igDen, 
    directed = T) 
      # where n is the number of nodes, 
      # p.or.m is the probability of drawing an edge, 
      #directed is whether the network is directed or not

#**plot(smugEL_DigRandoG, ...** in .R notebook  
     # vertex.size = 5, 
     # vertex.label = 1, 
     # edge.curved = .1, 
     # vertex.color = "tomato", 
     # edge.arrow.size = .1, 
     # edge.width = .5, 
     # edge.color = "grey60")

```{r iGraph: Directed: Degree Centrality}
# Degree centrality is calculated using the degree function in R. It returns how many edges each node has.
view(degree(ingTypeEL_Dig))

ingT_deg <- ingTypeEL_Dig
V(ingT_deg)$degree <- degree(ingT_deg) # assignment

# plot(ingT_deg, 
#     vertex.label.cex = .6, 
#     vertex.label.color = "black", 
#     vertex.size = V(ingTypeEL_Dig)$degree, 
#     vertex.label.cex = .2) # sized by degree

```{r iGraph: Directed: Betweenness Centrality}

view(betweenness(ingTypeEL_Dig, directed = FALSE))

# view(betweenness(ingTypeEL_Dig, directed = TRUE))
              # Yeilds 0

ingTypeEL_Between <- ingTypeEL_Dig
# We can again assign the output of betweenness() to a variable in the network and size the nodes according to it.

V(ingTypeEL_Between)$betweenness <- 
  betweenness(ingTypeEL_Between, 
              directed = F)

# plot(ingTypeEL_Between, 
#     vertex.label.cex = .6, 
#     vertex.label.color = "black", 
#     vertex.size = V(ingTypeEL_Between)$betweenness) # sized by betweenness

With closeness centrality we again make use of the shortest paths between nodes. We measure the distance between two nodes as the length of the shortest path between them. Farness, for a given node, is the average distance from that node to all other nodes. Closeness is then the reciprocal of farness (1/farness).
```{r iGraph: UnDirected:  Closeness Centrality}

view(closeness(smugEL_ig))

  # We assign it to a node variable and plot the network, adjusting node size by closeness.
ingTyp_Cent <- smugEL_ig #**UNDIRECTED **

V(ingTyp_Cent)$closeness <- closeness(ingTyp_Cent)

View(ingTyp_Cent)

# plot(ingTyp_Cent,
#     vertex.label.cex = .6, 
#     vertex.label.color = "black", 
#     vertex.size = V(ingTyp_Cent)$closeness/max(V(ingTyp_Cent)$closeness) * 20)

Degree centrality only takes into account the number of edges for each node, but it leaves out information about ego’s alters
```{r iGraph: Directed: Eigenvector Centrality}

view(evcent(ingTypeEL_Dig)$vector)
# view(evcent(smugEL_ig)$vector)
ingTyp_Vector <- ingTypeEL_Dig
V(ingTyp_Vector)$eigenvector <- evcent(ingTyp_Vector)$vector

# plot(ingTyp_Vector,
#     vertex.label.cex = .6, 
#     vertex.label.color = "black", 
#     vertex.size = V(ingTyp_Vector)$eigenvector/max(V(ingTyp_Vector)$eigenvector) * 20)
```

 “Interpretively, the Bonacich power measure corresponds to the notion that the power of a vertex is recursively defined by the sum of the power of its alters. The nature of the recursion involved is then controlled by the power exponent: positive values imply that vertices become more powerful as their alters become more powerful (as occurs in cooperative relations), while negative values imply that vertices become more powerful only as their alters become weaker (as occurs in competitive or antagonistic relations).”
```{r iGraph: Directed: Bonacich Centrality}

ingTyp_Cent <- ingTypeEL_Dig

V(ingTyp_Cent)$bonacich <- 
  power_centrality(ingTyp_Cent, exponent = -2, rescale = T)

V(ingTyp_Cent)$bonacich <- 
  ifelse(V(ingTyp_Cent)$bonacich < 0, 0, V(ingTyp_Cent)$bonacich)
```

```{r iGraph: Directed: Page Rank}

ingTyp_Cent <- ingTypeEL_Dig

V(ingTyp_Cent)$page_rank <- page_rank(ingTyp_Cent, directed = TRUE)$vector

 plot(ingTyp_Cent,
     vertex.label.cex = .6, 
     vertex.label.color = "black", 
     vertex.size = V(ingTyp_Cent)$page_rank/max(V(ingTyp_Cent)$page_rank) * 20)

```{r FUCK IT: Measure Correlations, eval=FALSE, include=FALSE}
# extract all the vertex attributes
all_atts <- 
  lapply(list.vertex.attributes(ingTyp_Cent),function(x) get.vertex.attribute(ingTyp_Cent,x))

# bind them into a matrix
all_atts <- 
  do.call("cbind", all_atts)

# add column nams
colnames(all_atts) <- 
  list.vertex.attributes(ingTyp_Cent)

# drop the family variable
all_atts <- 
  data.frame(all_atts[,2:ncol(all_atts)])

# convert all to numeric
all_atts <- 
  sapply(all_atts, as.numeric)

# produce a correlation matrix
cormat <- 
  cor(all_atts)

# melt it using reshape to function melt() to prepare it for ggplot which requires long form data
setDT(as.data.frame(cormat))
melted_cormat <- melt(cormat)

ggplot(data = melted_cormat, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile() +
  scale_fill_distiller(palette = "Spectral", direction=-2) +
  xlab("") +
  ylab("")
```

```{r FUCKIT : Centralization and Degree Distributions, eval=FALSE, include=FALSE}
# extract all the vertex attributes
all_atts <- lapply(list.vertex.attributes(ingTyp_Cent),function(x) get.vertex.attribute(ingTyp_Cent,x))
# bind them into a matrix
all_atts <- do.call("cbind", all_atts)
# add column nams
colnames(all_atts) <- list.vertex.attributes(ingTyp_Cent)
# drop the family variable
all_atts <- data.frame(all_atts[,2:ncol(all_atts)])
# convert all to numeric
all_atts <- sapply(all_atts, as.numeric)
# produce a correlation matrix
cormat <- cor(all_atts)

as.data.table(cormat)

view()
# melt it using reshape to function melt() to prepare it for ggplot which requires long form data
melted_cormat <- melt(cormat)
ggplot(data = melted_cormat, aes(x=Var1, y=Var2, fill=value)) + 
  geom_tile() +
  scale_fill_distiller(palette = "Spectral", direction=-2) +
  xlab("") +
  ylab("")
```
